### 2016-06-26 store-consistency.tex
Edits.

Reasoning under weakly consistent replication has received special
attention in the recent work~\cite{gotsmanpopl16}. Our operational
semantics and proof system are general enough to admit replication as
a special case. In this section, we explain how our reasoning approach
works under weakly consistent replication by comparing our model with
the standard model of replication.

Reasoning about eventually consistent replication often involves
reasoning in terms of \emph{replicas} of data. The primary challenge
in this setting is to ensure that the assumptions made and guarantees
enforced by an operation at one replica carry over to other replicas
that merge its effects, thus preserving the overall integrity of the
system. In~\cite{gotsmanpopl16}, this challenge is partly addressed by
imposing restrictions on how various replica states differ, i.e., by
fixing a system model with a stronger baseline consistency ({\sc cc})
than {\sc ec}. This unfortunately restricts the applicability of
reasoning approach from being extended to weaker stores such as
Bayou~\cite{bayou} or Quelea~\cite{pldi15}. Our view of replication
however does not explicitly involve replicas.
Fig.~\ref{fig:ec-theirs-vs-ours} contrasts our model of weakly
consistent replication with the conventional replica-based model.
Under our model, the notion of a replica is subsumed by the concept of
visibility; a replica is defined by the subset ($S$) of global state
($\E.\A$) that an operation witnesses. Constraints over replica states
therefore manifest as constraints over the visibility relation.  For
example, instead of requiring the store to be causally consistent, an
operation can request to witness a causally consistent subset of the
state; such demands can be made via the trace invariant $\I$. For a
precondition ($P$) of the operation to be useful, it has to be an
assertion over every causally consistent subset of the global state.
Since any replica that eventually executes the operation has to expose
one such subset ($S$), the precondition is guaranteed to hold
regardless of the replica. There is however one problem with this
explanation; by considering subsets of just one global state, it
ignores the fact that the global state (hence, the replica states)
change during the execution of the operation. To account for this
change,~\cite{gotsmanpopl16} distinguishes between \emph{effect
generation} event at one replica and \emph{effect merge} event at
another replica, and requires that certain invariants be preserved
between these two events at different replicas. Our framework folds
all of this into a stability condition (\S\ref{sec:rely-guarantee}).
Since any change to the global state during the execution of the
operation is an interference, and $P$ is required to be stable with
respect to any such interference, it follows that $P$ is valid on
every replica, thus ensuring that assumptions made at the generation
event are also valid at the merge event.

>>>>

Reasoning under weakly consistent replication has received special
attention in the recent work~\cite{gotsmanpopl16}. Our operational
semantics and proof system are general enough to admit replication as
a special case. In this section, we explain how the standard artifacts
of weakly consistent replication manifest in our reasoning framework.

Reasoning about weakly consistent replication often involves reasoning
in terms of \emph{replicas} of data. The primary challenge in this
setting is to ensure that the assumptions made and guarantees enforced
by an operation at one replica carry over to other replicas that merge
its effects, thus preserving the overall integrity of the system.
In~\cite{gotsmanpopl16}, this challenge is partly addressed by
imposing restrictions on how various replica states differ, i.e., by
fixing a system model with a stronger baseline consistency ({\sc cc})
than {\sc ec}. This unfortunately restricts the reasoning approach
from being applied to weaker stores such as Bayou~\cite{bayou} or
Quelea~\cite{pldi15}. Our view of replication however does not
explicitly involve replicas.  Fig.~\ref{fig:ec-theirs-vs-ours}
contrasts our model of weakly consistent replication with the
conventional replica-based model.  Under our model, the notion of a
replica is subsumed by the concept of visibility; a replica is defined
by the subset ($S$) of global state ($\E.\A$) that an operation
witnesses. Constraints over replica states therefore manifest as
constraints over the visibility relation.  For example, instead of
requiring the store to be causally consistent, an operation can
request to witness a causally consistent subset of the state; such
demands can be made via the trace invariant $\I$. For a precondition
($P$) of the operation to be useful, it has to be an assertion over
every causally consistent subset of the global state.  Since any
replica that eventually executes the operation has to expose one such
subset ($S$), the precondition is guaranteed to hold regardless of the
replica. There is however one problem with this explanation; by
considering subsets of just one global state, it ignores the fact that
the global state (hence, the replica states) change during the
execution of the operation. To account for this
change,~\cite{gotsmanpopl16} distinguishes between \emph{effect
generation} event at one replica and \emph{effect merge} event at
another replica, and requires that certain invariants be preserved
between these two events at different replicas. Our framework folds
all of this into a stability condition (\S\ref{sec:rely-guarantee}).
Since any change to the global state during the execution of the
operation is an interference, and $P$ is required to be stable with
respect to any such interference, it follows that $P$ is valid on
every replica, thus ensuring that assumptions made at the generation
event are also valid at the merge event.
